---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(2022)
```

# cartographer

<!-- badges: start -->
[![cartographer status badge](https://cidm-ph.r-universe.dev/badges/cartographer)](https://cidm-ph.r-universe.dev)
<!-- badges: end -->

Cartographer makes it easy to visualise your data on a map when you have a column
in the data that names a place/region. The map data itself is provided by other
packages or your own code.

## Installation

You can install cartographer like so:

``` r
options(repos = c(
  cidmph = 'https://cidm-ph.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
  
install.packages('cartographer')
```

## Map data

Packages can regiter map data with cartographer:

  * `{nswgeo}` - map data of New South Wales, Australia.
    
Alternatively, you can either register your own data using `register_map()`.

## Example

Cartographer comes with one example dataset registered. You can register additional
map datasets or depend on another package that registers them for you.

```{r}
library(ggplot2)
library(cartographer)

# list registered datasets
feature_types()

# list feature names in one of the datasets
head(feature_names("sf.nc"))
```

All you need is a column in your dataset that matches the `feature_names()` of
the map data. For the `"sf.nc"` map data, this dataset is suitable:

```{r}
head(nc_type_example)
```

You can then use this data with the geoms to provide map summaries. For the map
outline, you'll need to specify `feature_type`, but the other geoms will usually
be able to guess the correct value.

```{r example-basic, fig.width=8, fig.height=3}
ggplot(nc_type_example, aes(location = location)) +
  geom_boundaries(feature_type = "sf.nc") +
  geom_geoscatter(aes(colour = type), size = 0.5)
ggplot(nc_type_example, aes(location = location)) +
  geom_boundaries(feature_type = "sf.nc") +
  geom_centroids(aes(colour = type, scale = 6), size = 0.5)
ggplot(nc_type_example, aes(location = location)) +
  geom_boundaries(feature_type = "sf.nc") +
  geom_choropleth() +
  scale_fill_steps(low = "#e6f9ff", high = "#00394d", na.value = "white")
ggplot(nc_type_example, aes(location = location)) +
  geom_boundaries(feature_type = "sf.nc") +
  geom_pie(aes(fill = type), pie_radius = 0.1)
```

