---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(2022)
```

# cartographer

<!-- badges: start -->
[![cartographer status badge](https://cidm-ph.r-universe.dev/badges/cartographer)](https://cidm-ph.r-universe.dev)
<!-- badges: end -->

Cartographer makes it easy to visualise your data on a map when you have a column
in the data that names a place/region. The map data itself is provided by other
packages or your own code.

## Installation

You can install cartographer like so:

``` r
options(repos = c(
  cidmph = 'https://cidm-ph.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
  
install.packages('cartographer')
```

## Map data

Packages can register map data with cartographer:

  * `{nswgeo}` - map data of New South Wales, Australia.
    
Alternatively, you can either register your own data using `register_map()`.

## Example

Cartographer comes with one example dataset registered (the map of North Carolina
provided by the `{sf}` package where the feature names correspond to counties).
If you install the `{maps}` package, some additional maps are available.

```{r}
library(cartographer)

# list registered datasets
feature_types()

# list feature names in one of the datasets
head(feature_names("sf.nc"))
```

Our example dataset has a column called `county` with the name of the county and
a column `type` that is one of two values. We'd like to draw a map that shows
us where the entries are by type.

```{r}
head(nc_type_example_1)
```

We want to convert our data frame into a spatial data frame, then we'll be able to
use standard `{ggplot2}` tools to draw our map. Cartographer just needs to know
which of its maps we're using (`"sf.nc"`, although it's able to guess that if we
don't specify it), and which column has the feature names (`county`):

```{r, error = TRUE}
add_geometry(nc_type_example_1, county, feature_type = "sf.nc")
```

What happened? The matching of county names was case-insensitive, so that wasn't
the issue; it looks like our data contained a typo. One of the rows contains
`"PAMILCO"` instead of `"PAMLICO"` in the `county` column.

```{r}
library(dplyr, warn.conflicts = FALSE)

# cleaned dataset
nc_type_example_1_fixed <-
  nc_type_example_1 |>
  mutate(county = case_match(county, "PAMILCO" ~ "PAMLICO", .default = county))

add_geometry(nc_type_example_1_fixed, county, feature_type = "sf.nc")
```

Now that we have our spatial geometry added, we can go ahead and plot it, adding
a layer that uses the full `"sf.nc"` dataset to give us the base map:

```{r example-basic, fig.width=8, fig.height=2}
library(ggplot2)

nc_type_example_1_fixed |>
  count(county, type) |>
  add_geometry(county, feature_type = "sf.nc") |>
  ggplot() +
  geom_sf(data = map_sf("sf.nc")) +
  geom_sf(aes(fill = n)) +
  facet_wrap(vars(type))
```

The package `{ggautomap}` uses cartographer to provide some new geoms that
transparently attach the map data and perform some calculations like mapping our
data to the middle or random points within the corresponding feature.
