---
title: "Registering map data with 'cartographer'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Registering map data with 'cartographer'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cartographer)
```


If you are often working with spatial data concerning a specific region, it might
make sense to bundle up the map data along with its registration as an R package.
This is true even if the data is only for internal use.

For a general introduction to R packaging, see [the R packaging book](https://r-pkgs.org/).
The [chapter on packaging data](https://r-pkgs.org/data.html) in particular contains many
relevant hints.


## General guidelines
Your source dataset needs to be transformed into a `{sf}` data frame. The `sf::read_sf()`
can ingest many popular spatial data formats
Make use of `usethis::use_data_raw()` and `usethis::use_data()` to prepare the data.

Your package will need to inform `{cartographer}` of the data by calling `register_map()`.
This needs to happen after your package is loaded, so the `base::.onLoad()` hook is the
appropriate place:

```{r}
# in R/zzz.R
.onLoad <- function(libname, pkgname) {
   cartographer::register_map(
     "my_package.uk",
     data = rnaturalearth::ne_states(country = "united kingdom", returnclass = "sf"),
     feature_column = "name_en",
     outline = rnaturalearth::ne_countries(country = "united kingdom", returnclass = "sf", scale = "large")
   )
}
```

The hook needs to come last when your package is evaluated, so the convention is
to include it in a file called `R/zzz.R`, which is normally alphabetically last.
The registration should at least include the map data and the name of the column that
labels the features. It can also contain an alias mapping if there are alternative
names or abbreviations for places that would plausibly appear in datasets.


## Reducing file size
Spatial datasets can often be quite large as they contain high resolution data.
This can especially be a problem if attempting to commit the data to git or publish
your package to CRAN (where 1 MB is the soft limit for the total compressed size of
packages).

Consider removing extra columns from the spatial dataset so that only the geometry column
and the name column remain.

A useful pattern to reduce the size is to simplify the geometry to a specified resolution:

```{r, eval=FALSE}
# Preserve the original coordinate reference system.
crs_orig <- sf::st_crs(high_res_sf_data_frame)

# Convert to a more suitable CRS for manipulation. Note that the lat_ts argument
# here is the "latitude of true scale", i.e. the latitude at which scale will be
# the least distorted. Adjust this based on your data.
crs_working <- sf::st_crs("+proj=eqc +lat_ts=34 units=m")

# Choose a resolution: features smaller than this scale will be lost.
tolerance_m <- 1000L

low_res <- high_res_sf_data_frame |>
  sf::st_transform(crs_working) |>
  sf::st_simplify(dTolerance = tolerance_m) |>
  sf::st_transform(crs_orig)

# Compare the size after reducing the resolution:
object.size(high_res_sf_data_frame)
object.size(low_res)
```

See the documentation for the [equidistant cylindrical](https://proj.org/operations/projections/eqc.html)
projection for details of its configuration.

You may also wish to remove holes in features:

```{r, eval=FALSE}
new_geom <- geom |>
  sf::st_transform(crs_working) |>
  sf::st_union() |>
  nngeo::st_remove_holes() |>
  sf::st_make_valid()
```

